---
title: "Practical Machine Learning - Week 4 Assignment"
output: html_document
---


# Prediction of correct activity movement


This report contains a prediction model on activity data (barbell lifts) from wearables. The goal of the model is to predict the outcome whether the activity is correct/incorrect.


## Load and clean data

The datasets are loaded and columns containing NA's or non-relevant information are removed

```{r}
library(readr)

# Load training/test set data
test <- read_csv("pml-testing.csv")
train <- read_csv("pml-training.csv")


# Remove NA variables from test and trainig set
train <- train[, colSums(is.na(test)) != nrow(test)]
test <- test[, colSums(is.na(test)) != nrow(test)]

# Remove names/ timestamp / window related columns
train  <-  subset(train, select = -c(1,2,3,4,5,6,7))
test  <-  subset(test, select = -c(1,2,3,4,5,6,7)) 

# Split variables and outcomes
train_Y <- subset(train, select = c(53))

train_Y$classe <- as.factor(train_Y$classe)

train_X <- subset(train, select = -c(53))

test_X <- subset(test, select = -c(53))

```

We look into the dataset in more detail.

```{r}
# Dimensions of training data set
dim(train_X)

# Number of unique outcomes
unique(train_Y$classe)

```
## Random Forest

We use a random forest model with 10-fold cross validation on the training set.

```{r}
library(caret)
library(randomForest)

set.seed(1)


# Use 10-fold cross validation
settings <- trainControl(method="cv", number=10, verboseIter=FALSE)

rfModel <- randomForest(x = train_X, y = train_Y$classe,
                        method = "rf",
                        trControl = settings)

```

We check the accuracy of the trained model against the training set.

```{r}

predictRF <- predict(rfModel, newdata=train_X)
confMatRF <- confusionMatrix(predictRF, train_Y$classe)
confMatRF

plot(confMatRF$table, col = confMatRF$byClass, 
     main = paste("RF- Accuracy =", round(confMatRF$overall['Accuracy'], 2)))


```



We now predict the test set:
```{r}

predictRF_test <- predict(rfModel, newdata=test_X)
 
head(predictRF_test,20)

```

## Note:

The random forest model gives a very good accuracy at the expense of interpretability of the model.
Considering the large amount of variables, it may be a good approach to also check other model approaches that reduce the amount of (correlated) variables in the model. Such as regression over principal components, or PLSDA.
